---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd -->

# qarPI

<!-- badges: start -->
<!-- Si activas GitHub Actions con usethis::use_github_action_check_standard(),
     descomenta el badge de R-CMD-check y ajusta USUARIO/REPO -->
 <!--[![R-CMD-check](https://github.com/SilviaNovo/qarPI/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/SilviaNovo/qarPI/actions/workflows/R-CMD-check.yaml) -->
 [![License: GPL v3](https://img.shields.io/badge/License-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
 [![arXiv](https://img.shields.io/badge/arXiv-2512.22018-b31b1b.svg)](https://doi.org/10.48550/arXiv.2512.22018)
<!-- badges: end -->


The package provides prediction intervals (PIs) under two model classes:

- Classical autoregressive models: **AR(p)**
- Quantile autoregressive models: **QAR(p)**

A central feature of the package is the implementation of the new bootstrap algorithms 
**AR-perc**, **AR-proot**, **QAR-perc**, and **QAR-proot**, introduced in the article 
*тАЬPrediction Intervals for Quantile AutoregressionтАЭ* ([arXiv:2512.22018](https://doi.org/10.48550/arXiv.2512.22018)). 

In addition, the package includes several established bootstrap and analytical procedures (Box-Jenkins)
from the existing literature, allowing users to replicate and compare alternative 
PI constructions within a unified framework.

Within each model class, bootstrap PIs may therefore be obtained via:

- Percentile-based bootstrap procedures
- Predictive-root-based bootstrap procedures


---


### Installation

```{r,results='hide',message=FALSE,warning=FALSE}
# install the package
devtools::install_github("SilviaNovo/qarPI")


#load the package
library(qarPI)
```
---  


### 1. PI methods for AR(p) models


#### Percentile-based bootstrap  PIs

Bootstrap procedures constructed under the AR(p) framework using the percentile principle

- **AR-perc** тАФ Bootstrap algorithm based on quantile estimation and bootstrap multipliers (Novo & Sanchez-Sellero, 2025)
  тЖТ `pi_AR_perc()`
  
```{r}

set.seed(123)
y <- arima.sim(model = list(ar = 0.6), n = 300)

# AR-perc PI
pi <- pi_AR_perc(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
```

- **TS** тАФ Backward bootstrap (Thombs & Schucany, 1990)  
  тЖТ `pi_TS()`
  
```{r}
# Thombs & Schucany PI
pi <- pi_TS(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
```

- **CB** тАФ Conditional bootstrap (Cao et al., 1997)  
  тЖТ `pi_CB()`
  
```{r}
# Cao et al. PI
pi <- pi_CB(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
``` 

- **PRR** тАФ Forward bootstrap  (Pascual, Romo & Ruiz, 2004) using ordinary least squares (OLS) or least absolute deviations (LAD) estimation 
  тЖТ `pi_PRR()`
  
```{r}
# Pascual et al. PI. OLS
pi <- pi_PRR(y, p = 1, h = 3, method="OLS")

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI

# Pascual et al. PI. LAD
pi <- pi_PRR(y, p = 1, h = 3, method="LAD")

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI

``` 



---



#### Predictive-root-based bootstrap PIs

Bootstrap intervals constructed using the predictive root approach

- **AR-proot** тАФ Bootstrap algorithm based on quantile estimation, bootstrap multipliers and predictive residuals (Novo & Sanchez-Sellero, 2025)
  тЖТ `pi_AR_proot()`

```{r}
# AR-proot PI
pi <- pi_AR_proot(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
``` 


- **PP** тАФ Forward bootstrap with predictive residuals  (Pan & Politis, 2016, denoted by the 
authors as *Fp*). 
  тЖТ `pi_PP()`
  
```{r}
# Pan & Politis PI
pi <- pi_PP(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
``` 
  
  

#### Analytical PIs: Box-Jenkins

- **BJ** тАФ BoxтАУJenkins type analytical PIs assuming Gaussian innovations  
  тЖТ `pi_BJ()`
  
```{r}

# Classical BoxтАУJenkins PI
pi <- pi_BJ(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
```
---

---

### 2. PI methods for QAR(p) models

Prediction intervals constructed under the QAR(p) framework

---

#### Percentile-based bootstrap PIs

- **QAR-perc** тАФ Bootstrap algorithm based on quantile estimation and bootstrap multipliers (Novo & Sanchez-Sellero, 2025)
  тЖТ `pi_QAR_perc()`

```{r}
# QAR-perc PI
pi <- pi_QAR_perc(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
```  

- **X** тАФ Bootstrap algorithm suggested by Xiao (2012)  
  тЖТ `pi_X()`
  
```{r}
# Xiao PI
pi <- pi_X(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
``` 
  



---

#### Predictive-root-based bootstrap PIs

- **QAR-proot** тАФ Bootstrap algorithm based on quantile estimation, bootstrap multipliers and predictive residuals (Novo & Sanchez-Sellero, 2025)
  тЖТ `pi_QAR_proot()`
  
```{r}
# QAR-proot PI
pi <- pi_QAR_proot(y, p = 1, h = 3)

t_PI <- data.frame(
  Lower = pi$lpi,
  Upper = pi$upi,
  Length = pi$len
)
t_PI
``` 

---


## Summary Structure

```text
qarPI
тФВ
тФЬтФАтФА AR(p)
тФВ   тФЬтФАтФА Percentile-based Bootstrap
тФВ   тФВ   тФЬтФАтФА AR-perc тЖТ pi_AR_perc()
тФВ   тФВ   тФЬтФАтФА TS тЖТ pi_TS()
тФВ   тФВ   тФЬтФАтФА CB тЖТ pi_CB()
тФВ   тФВ   тФФтФАтФА PRR тЖТ pi_PRR()
тФВ   тФВ
тФВ   тФЬтФАтФА Predictive-root-based Bootstrap
тФВ   тФВ   тФЬтФАтФА AR-proot тЖТ pi_AR_proot()
тФВ   тФВ   тФФтФАтФА PP (Fp) тЖТ pi_PP()
тФВ   тФВ
тФВ   тФФтФАтФА Analytical
тФВ       тФФтФАтФА BJ тЖТ pi_BJ()
тФВ
тФФтФАтФА QAR(p)
    тФЬтФАтФА Percentile-based Bootstrap
    тФВ   тФЬтФАтФА QAR-perc тЖТ pi_QAR_perc()
    тФВ   тФФтФАтФА X тЖТ pi_X()
    тФВ
    тФФтФАтФА Predictive-root-based Bootstrap
        тФФтФАтФА QAR-proot тЖТ pi_QAR_proot()
```

## References

- Box, G. E. & Jenkins, G. M. (1976). *Time Series Analysis: Forecasting and Control*. HoldenтАУDay, San Francisco.

- Cao, R., et al. (1997). Saving computer time in constructing consistent bootstrap prediction intervals for autoregressive
processes. *Communications in Statistics-Simulation and Computation*, 26(3):961тАУ978. https://doi.org/10.1080/03610919708813420

- Novo, S., & S├бnchez-Sellero, C. A. Prediction intervals for quantile autoregression. https://doi.org/10.48550/arXiv.2512.22018


- Pan, L., & Politis, D. N. (2016). Bootstrap prediction intervals for linear, nonlinear and nonparametric autoregressions. *Journal of Statistical Planning and Inference*. https://doi.org/10.1016/j.jspi.2014.10.003

- Pascual, L., Romo, J., & Ruiz, E. (2004). Bootstrap predictive inference for ARIMA processes. *Journal of Time Series Analysis*. 25(4):449тАУ465. https://doi.org/10.1111/j.1467-9892.2004.01713.x


- Thombs, L. A., & Schucany, W. R. (1990). Bootstrap prediction intervals for autoregression. *Journal of the American Statistical Association*, 85(410):486тАУ492. https://doi.org/10.1080/01621459.1990.10476225

- Xiao, Z. (2012). Time series quantile regressions. Handbook of Statistics. Time Series Analysis: Methods and Applications, 30:213тАУ257. https://doi.org/10.1016/B978-0-444-53858-1.00009-0
